---
title: "Reproducible Research: Peer Assessment 1"
author: Natsuno Aoki
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
# 1. Load the data
```{r}
dat <- read.csv("activity.csv")
```
# 2. Converting variables into date-class and numeric-class
```{r}
dat$date <- as.Date(dat$date, "%Y-%m-%d")
dat$interval <- as.numeric(dat$interval)
dat$steps <- as.numeric(as.character(dat$steps))
```

## What is mean total number of steps taken per day?
# 1. Calculate the total number per date
```{r}
library(plyr)
library(dplyr)
sumd <- dat %>%
        group_by(date) %>%
        summarise(total = sum(steps, na.rm = TRUE))
head(sumd, type = "html")
```
# 2. Make a histogram for total numbers per day
```{r}
library(ggplot2)
ggplot(sumd, aes(x=date, y=total)) + 
        geom_bar(stat = "identity") +
        labs(x = "Date", y = "Total numbers of steps", title = "Total numbers per day")
```

# 3. Calculate the mean and meadian of the total number of steps per day
 - mean
```{r}
mean(sumd$total)
```
 - median
```{r}
median(sumd$total)
```
## What is the average daily activity pattern?
# 1. Make a time series plot
 - Prepare the mean data per interval
```{r}
detach(package:plyr)
meand <- dat %>%
        group_by(interval) %>%
        summarise(mean = mean(steps, na.rm = TRUE))
head(meand, type = "html")
```
 - Time series plot of the average number of steps
```{r}
ggplot(meand, aes(x=interval, y=mean)) + 
        geom_line(stat = "identity") +
        labs(x = "Interval", y = "Average numbers of steps", title = "Time series plot of the average number of steps")
```

# 2. Which 5-minute interval contains the maximum number of steps?
```{r}
meand$interval[meand$mean == max(meand$mean)]

```

## Imputing missing values
# 1. Calculate the total number of rows with NAs
```{r}
nrow(subset(dat, is.na(dat$steps)))
```
# 2-3. Replace NAs with the mean value of the corresponding intervals
 - If there is a NA in the column "steps", replace it with the mean value. Otherwise, remain the original value in the steps.
```{r}
dat_mean <- dat %>%
        group_by(interval) %>%
        mutate(steps.mean = ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps))

head(dat_mean, type = "html")
```
 - In the case of replacing NAs with the medians
```{r}
dat_med <- dat %>%
        group_by(interval) %>%
        mutate(steps.median = ifelse(is.na(steps), median(steps, na.rm = TRUE), steps))

head(dat_med, type = "html")
```
4. Make a histogram of the total number per day
 - Prepare the total data using the filled-in dataset with the median
```{r}
sum_dat_med <- dat_med %>%
        group_by(date) %>%
        summarise(total = sum(steps.median, na.rm = TRUE))

head(sum_dat_med)
```

- Using the filled-in dataset with the median
```{r}
sum_dat_mean <- dat_mean %>%
        group_by(date) %>%
        summarise(total = sum(steps.mean, na.rm = TRUE))

head(sum_dat_mean)
```

 - Make a histogram
```{r}
ggplot(sum_dat_med, aes(x=date, y=total)) + 
        geom_bar(stat = "identity", colour = "white") +
        labs(x = "Date", y = "Total numbers of steps", title = "Total numbers per day")
```

3 plots for comparison
```{r}
#Make a new column for distincting original data/filled-in data
sumd$filled <- c("original")
sum_dat_med$filled <- c("median")
sum_dat_mean$filled <- c("mean")
alld <- rbind(sumd, sum_dat_med, sum_dat_mean)

g <- ggplot(alld, aes(x = date, y = total))
g + geom_line(aes(colour = filled)) + labs(x = "Date", y = "Total of steps", title = "Comparison of the original data with NA / without NA")
```

 - The mean total number of steps per day, using the filled-in data
```{r}
#Original data
mean(sumd$total)

#NA-filled with the median
mean(dat_med$steps.median)

#NA-filled with the mean
mean(dat_mean$steps.mean)
```

 - The median total number of steps per day, using the filled-in data
```{r}
#Original data
median(sumd$total)

#NA-filled with the median
median(dat_med$steps.median)

#NA-filled with the mean
median(dat_mean$steps.mean)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
#Using the data with NAs replaced with the mean
Sys.setlocale("LC_TIME","en_US")
dat_w <- dat_mean %>%
        mutate(week = ifelse(weekdays(date, abbreviate = FALSE) == "Saturday", "weekend", ifelse(weekdays(date, abbreviate = FALSE) == "Sunday", "weekend","weekday")))
```

 - Mean over the weekdays
```{r}
wd <- subset(dat_w, week == "weekday")
wd <- select(wd, steps.mean, week, interval)
wd_av <- wd %>%
        group_by(interval) %>%
        mutate(mean = mean(steps.mean, na.rm = TRUE))

head(wd_av)
```
 
 - Mean over the weekend
```{r}
we <- subset(dat_w, week == "weekend")
we <- select(we, steps.mean, week, interval)
we_av <- we %>%
        group_by(interval) %>%
        mutate(mean = mean(steps.mean, na.rm = TRUE))

head(we_av)
```
 
 - Make 2 plots: average over weekdays/weekend
```{r}
weekdata <- rbind(wd_av, we_av)
 #Set the window
g <- ggplot(weekdata, aes(x = interval, y = mean))
g + geom_line() + ylim(0,200) + facet_grid(.~week)
```
